Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index Using Machine Learning Methods

### 2024-08-13 tex/paper.tex
updated references

\begin{abstract}
    Investment decisions and economic policy are highly influenced by 
    macroeconomic variables and the stock market. Econometric models
    have been the predominant tools in this context. The aim of this study is to analyse the impacts of some especially important macroeconomic
    indicators on the FTSE250 
    (Financial Times Stock Exchange 250 Index), using monthly data 
    from  October 1993 to April 2024 using an ARDL model, obtaining 
    quantifiable estimates for the long-term and short-term impacts of the variables
    on the FTSE250. The efficacy of Radial Basis Function 
    Neural Networks (RBF-NN) in this problem domain is then investigated. We find that 
    the RBF-NN is a competent short-term forecasting tool, and is able to 
    provide an alternative approach to standard econometric models in measuring the long-term influence of
    macroeconomic variables on the stock market.
\end{abstract}

@@

In this study we first use an Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM), to investigate 
the different effects of selected macroeconomic variables on the FTSE250 Index over the January 1993 - March 2024 period.
Second, we test the efficacy of a Radial Basis Function Neural Network (RBF-NN) in forecasting the price movements of the FTSE250 based on the selected macroeconomic indicators, over the short-term 
(5 months) and long-term (57 months).

We utilise the FTSE 250 as our measure of the UK stock market, because unlike the 
FTSE100 its overseas share of earnings is 
roughly $50\%$, similar to the S\&P 500. In contrast, nearly $80\%$ of the FTSE $100$'s earnings 
are from overseas. This is reflected in the constitution of the respective indices, with 
the FTSE 250 historically resembling the structure 
of the UK economy more closely than the FTSE100.

The macroeconomic variables we consider are inflation, indicated by the Consumer Price
Index (CPI); interest rate, represented by the Bank of England policy rate; the exchange rate, 
represented by the USD/GBP exchange rate; and the money supply, represented by the 
M3 measure of money supply, preferred over other 
measures of the money supply because it considers more assets than M2 money supply 
but does not include extremely illiquid assets as the M4 measure does. The selection 
of these variables was based on the literature and the structure of the UK economy since 
the establishment of the FTSE250 Index.

@@

There are a number of works which investigate the impact of macroeconomic variables
on the stock market.

@@

Radial Basis Function Neural Networks (RBF-NN) are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network. They have since been employed in a variety of contexts such as 
hydrological forecasting \citep{chang2001}, biomedical engineering 
\citep{ibrikci2002} and time-series analysis \citep{xiong2015} with success. 
Whilst neural networks trained on small datasets often display unstable 
behaviour during performance \citep{lebaron1998}, The advantage of 
RBF-NNs lies in their ability to 
operate effectively in extremely volatile financial time-series 
environments \citep{cafferata2019}, their strong ability to 
generalise on such data \citep{sharkawy2020} as well as their efficacy on small datasets \citep{kosarac2022}. 
\citeboth{esfandyari2016}
supports this, using a sample size of 70 and 29 to forecast daily NASDAQ 
index values using an Artificial Neural Network; they achieved a minimum
test $R^2$ of 83$\%$. RBF-NNs have been used 
considerably for stock price forecasting. \citeboth{cao2004} 
used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
forecast S$\&$P 500 prices. \citeboth{dass2019} use an RBF-NN, with the Back-propagation 
Algorithm (BPA) used instead to tune parameters; their results demonstrate 
that the RBF-NN outperformed a standard feed-forward neural network. 
\citeboth{ji2014} also found that the RBF-NN, on a 
small training dataset, was able to predict the value 
of the Shanghai Composite Index on a 45 day time-horizon, with a mean prediction error of 
1.4387$\%$, suggesting that the RBF-NN is competent in time-series 
forecasting using small training datasets. Recently, \citeboth{abotaleb2024}
used quarterly data from 2001 to 2023 to model the influence of economic 
policies on the quarterly prices of various stock markets including the 
FTSE100 using an RBF-NN, finding that the model 
achieved relative forecast error rates as low as 6.8$\%$ and an error of 
27.2$\%$ in the case of the FTSE100. Thus, we find that the RBF-NN is a 
competent tool in the sphere of financial time-series forecasting, and 
is able to operate effectively within the constraints placed by our 
problem domain and data.

@@

Based on the empirical research and characteristics of the UK economy, we 
define the stock market index function as:

\begin{equation}
    FTSE_{250} = f\biggl(CPI, EXCHG, INT, M3\biggr)
\end{equation}

@@

Before constructing the ARDL model we first apply a natural logarithmic transform
on the variables, this is done because in a log-transformed model,
the coefficients can be interpreted as elasticities. Put simply, 
a coefficient in a model, where both the dependent and independent variables
are in logarithmic form, represents the percentage change in the 
dependent variable for a one percent change in an independent variable, 
which furnishes us with easily interpretable results. Having done this,
we must then check at which level the variables are stationary (sidenote) at.
Stationarity refers to the statistical property of a time-series where 
its mean, variance, and autocorrelation structure do not change over time; 
either a variable is stationary at level ($I(0)$), requiring no alteration,
or it is stationary after taking the first difference ($I(1)$). 
Non-stationarity leads to spurious regression, where relationships that 
are not actually present between two time-series are produced. Various
techniques exist for stationarity testing, however in this study the 
Augmented Dickey-Fuller (ADF) unit root test, developed by (link), was used.

As the ARDL model involves lagged values of the dependent and independent 
variables, we are required to determine the optimum lag for the variables 
that minimises error; this was done using the Schwarz
Information Criterion.

The implicit form (2) is represented in the ARDL model as:

@@

for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic is computed for the joint
significance of the lagged variables and compared against the critical values
given by Pesaran et. al. (2001) to determine whether cointegration exists, 
and at what significance level. If cointegration exists, then we estimate 
the long-run coefficients $\kappa_i$ by estimating the cointegration equation:

\begin{align}
    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon 
\end{align}

@@

where the error correction term ECT is found by calculating the residual 
produced by (3):

@@

The parameter $\gamma$ describes the speed at which short-run deviations from the long-run equilibrium
are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that converges the deviation 
to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
well as the short-run correction speed $\gamma$ we use OLS regression.

@@

We began by applying a natural logarithmic transform on the variables, and 
then conducting the ADF unit root test on the data. The null hypothesis of the 
ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
is that the variables are stationary.

@@

Subsequently, the optimum lag of the variables was calculated using the 
Schwarz Information Criterion, to obtain the optimum model 
of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.0798$ was compared against the critical values
given by Pesaran et. al. (2001), providing evidence of cointegration 
at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
level.

@@

\begin{table}[h!]
    \centering
    \caption{Coefficients Estimates}
    \begin{tabular}{llll}
        \toprule
        \textbf{Variable} & \textbf{Long-Run Coefficient} & \textbf{Variable} & \textbf{Short-Run Coefficient} \\
        \midrule
        FTSE$_{250}$ & N/A & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
        CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
        INT & $-0.1341097^{***}$ & INT $(-7)$ & $-0.034572^{****}$\\
        EXCHG &  $-0.890146^{****}$ & EXCHG $(-7)$ & $0.122318$ \\
        M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.16^{**}$ \\
        \bottomrule
    \end{tabular}
\end{table}

@@

The long term coefficients show that the interest rate and exchange rate
interest rate have significant long term effects on the stock market price
while money supply and inflation are statistically insignificant. Neifar (2023) 
proposes that this is because deflation, that is the negative movements in the 
CPI, is a statistically significant factor in the movement of the UK stock 
market, but positive movements (inflation) are not. Nevertheless, the positive
coefficient of 3.16, and the negative short-term adjustment, is in line with the expectation from the literature that 
inflation exerts a positive effect on the stock market in the long-term, but a negative 
one in the short-run.

The insignificant long-term effect of the money supply 
was noted by Picha (2017) and appears to be confirmed here. There are a few explanations
for why this may be.

Unsurprisingly, the USD/GBP exchange rate is significantly related to the FTSE 250's movements. It 
should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
at times $50\%$ of the FTSE 250's earnings have come from abroad, indicating that 
the value of the Pound has an important impact on this stock market. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it by $1\%$ has an impact of $-0.89\%$ on the FTSE 250. The findings 
are in line with Neifar (2023), Liang and Willett (2015), and Khan and Khan (2018).

@@

Turning to the short-run dynamics, we find that the error correction term 
is negative and statistically significant ensuring that long-run equilibrium 
can be achieved. The adjustment speed of $0.16$ indicates that 
roughly $16\%$ of the fluctuation in the previous month's price, caused by the 
macroeconomic variable movements, is adjusted back to the long-run equilibrium in the current month.
However, only increases in the interest rate and money supply are able to 
significantly contribute towards this correction, with $1\%$ increases in the former 
leading to a $-0.03\%$ change in the FTSE250 price 7-months later, and in the case of the latter
a $-0.53\%$ change after a 4-month lag.

@@

By introducing several modifications to the exact interpolation 
procedure, we obtain the RBF-NN
model first constructed in (Broomhead and Lowe, 1988). 
The RBF-NN mapping takes the following form:

@@

The RBF-NN was utilised for both short-term and long-term 
predictions. For short-term forecasting, 
the model was tested over a 5-month horizon, based on the optimum lag calculated earlier, while the long-term predictions spanned 57 months. 
We differenced the data based on the recommendations of the 
unit root tests conducted earlier to achieve stationarity.

We began by normalising the data, using a z-score normalisation. This is because
neural networks work best with data
that is between a specified range; it ensures that the data is roughly uniformly
distributed between the network inputs and outputs 
(Mendelsohn, 1993, Preprocessing Data for Neural Networks). Following this 
we had to program the RBF-NN into Python. As no specific RBF package exists
in Python, we used the kernel regression approach outlined by 
Bishop, using the GaussianProcessRegressor (GPR) package in Python.

@@

As this is a Bayesian model we need to specify a \textit{prior distribution} over the parameters, 
expressing our beliefs about the prior
parameters before we examine the observations. A covariance function or kernel
is specified according to the characteristics of the data; in our model this 
is the generalised inverse multiquadric

@@

\begin{align}
    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr)
\end{align}
where the covariance matrix $A^{-1} = K^{-1} + \frac{XX^T}{\sigma_{n}^2}$.
To make predictions on a test case all possible $w$
values, weighted by their posterior probability, are averaged to obtain the
predictive distribution, upon which predicted outputs are computed.

@@

A $4$-dimensional input space $X$
passes inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \), through the first
layer, in which the covariance function $(x,x')$
is applied, the goal being to determine the distribution of the weights. Prior to this,
the optimal hyperparameters in the covariance function, which plays an analogous
role to the RBF basis function, are calculated using GridSearch Cross Validation; these hyperparameters are:
\begin{enumerate}
    \item $\alpha$: a regularisation parameter which represents a small quantity
    that is added to the diagonals of the covariance matrix to ensure it is 
    invertible. 
    \item $\beta$: the power in the covariance function, which controls the shape of the function.
    \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for; naturally, 
    it moderates the risk of overfitting and underfitting.
    \item Length scale: a parameter that controls the influence - analogous to 
    the width each RBF neuron possesses - that each kernel (analogous to node) is to have.
\end{enumerate}
following this the model uses the training observation $(y,X)$ to find
the posterior predictive distribution, and with it the predicted output, after 
averaging over all possible $w$, weighted by their likelihood according to (10).

@@

The interpretation of this is that, historically, the variations in the 
Consumer Price Index, interest rate, USD/GBP exchange rate, and M3 money 
supply have explained or accounted for $9\%$ of the fluctuations of the
FTSE250 value; in the context of a phenomenon as complex as the stock market,
which is influenced by an inumerable quantity of factors such as investor sentiment
- inherently irrational, 
political instability, geopolitical shocks, corporate governance, and financial regulation, this is 
unsurprising. Furthermore, the selection of economic variables in this study
is by no means exhaustive; commodity prices, international trade dynamics,
economic growth, and many other variables also exert a non-insignificant 
effect on the stock market.

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$, for a 5-month time horizon.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possesses
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price.

In the short-term, which we define as 5 months, 
as per the optimum lag for the FTSE250 in the ARDL model, the market movements
are more closely tied to recent deviations in the economy, because theoretically
the short-run is not exposed to the various unforeseen external shocks and structural changes
that dilute the relationship between these variables and the stock market in the 
long-run. Essentially, as we stretch our prediction window, we 
introduce more uncertainty and unknowns, which translates into a 
lower predictive quality.

@@

Furthermore, as the time horizon extends into the long-run, fundamental factors 
begin to matter a lot more. The compound effect of earnings, cash flows and
managerial efficiencies begins to outweigh the 
influence of fluctuating macroeconomic sentiment. Thus, in the long-term
we see a different picture altogether.

@@

@@

@@

The the study also investigated the efficacy of machine learning methods in 
the field of stock-econometric forecasting, and returned good results, providing investors
with an accurate short-term forecast tool, and also providing economists
with a quantitative non-econometric measure of the 
long-term influence of the selected macroeconomic 
variables on the FTSE250 that outperforms a standard econometric model.

Further research should endeavour to include other macroeconomic variables such as 
foreign direct and portfolio investment, commodity prices and economic growth.
Falling outside the immediate scope of the current study, 
an autoregresive model could also be tried, including the past values of the 
FTSE250 in the input space to provide a more accurate model. In addition, 
a more granular data set could be utilised by applying a 
Brownian bridge algorithm to generate daily macroeconomic data and 
subsequently testing the RBF-NN against other neural networks 
such as auto-regressive LSTMs and GRUs, as well as other machine learning methods suited for 
forecasting such as Support Vector Machines (SVM).

>>>>

\begin{abstract}
    Investment decisions and economic policy are highly influenced by 
    macroeconomic variables and the stock market. Econometric models
    have been the predominant tools in this context. The aim of this study is to analyse the impacts of some especially important macroeconomic
    indicators on the FTSE250 
    (Financial Times Stock Exchange 250 Index), using monthly data 
    from  October 1993 to April 2024 using an ARDL model, obtaining 
    quantifiable estimates for the long-term and short-term impacts of the variables
    on the FTSE250. The efficacy of Radial Basis Function 
    Neural Networks (RBF-NN) in this problem domain is then investigated. It is established 
    the RBF-NN is a competent short-term forecasting tool, and is able to 
    provide an alternative approach to standard econometric models in measuring the long-term influence of
    macroeconomic variables on the stock market.
\end{abstract}

@@

In this study, first an Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM), is employed to investigate 
the different effects of selected macroeconomic variables on the FTSE250 Index over the January 1993 - March 2024 period.
Second, the efficacy of a Radial Basis Function Neural Network (RBF-NN) is tested in forecasting the price movements of the FTSE250 based on the selected macroeconomic indicators, over the short-term 
(5 months) and long-term (57 months), the former intended to inform investors and the latter to provide
economic policymakers with a tool to measure the influence of macroeconomic variables on the price of the stock market.

The FTSE 250 is utilised as the measure of the UK stock market, because unlike the 
FTSE100 its overseas share of earnings is 
roughly $50\%$, similar to the S\&P 500. In contrast, nearly $80\%$ of the FTSE $100$'s earnings 
are from overseas. This is reflected in the constitution of the respective indices, with 
the FTSE 250 historically resembling the structure 
of the UK economy more closely than the FTSE100.

The macroeconomic variables considered are inflation, indicated by the Consumer Price
Index (CPI); interest rate, represented by the Bank of England policy rate; the exchange rate, 
represented by the USD/GBP exchange rate; and the money supply, represented by the 
M3 measure of money supply, preferred over other 
measures of the money supply because it considers more assets than M2 money supply 
but does not include extremely illiquid assets as the M4 measure does. The selection 
of these variables was based on the literature and the structure of the UK economy since 
the establishment of the FTSE250 Index.

@@

There are a number of works which investigate the impact of 
the selected macroeconomic variables
on the stock market.

@@

Radial Basis Function Neural Networks (RBF-NN) are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network. They have since been employed in a variety of contexts such as 
hydrological forecasting \citep{chang2001}, biomedical engineering 
\citep{ibrikci2002} and time-series analysis \citep{xiong2015} with success. 
Whilst neural networks trained on small datasets often display unstable 
behaviour during performance \citep{lebaron1998}, The advantage of 
RBF-NNs lies in their ability to 
operate effectively in extremely volatile financial time-series 
environments \citep{cafferata2019}, their strong ability to 
generalise on such data \citep{sharkawy2020} as well as their efficacy on small datasets \citep{kosarac2022}. 
\citeboth{esfandyari2016}
supports this, using a sample size of 70 and 29 to forecast daily NASDAQ 
index values using an Artificial Neural Network; they achieved a minimum
test $R^2$ of 83$\%$. RBF-NNs have been used 
considerably for stock price forecasting. \citeboth{cao2004} 
used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
forecast S$\&$P 500 prices. \citeboth{dass2019} use an RBF-NN, with the Back-propagation 
Algorithm (BPA) used instead to tune parameters; their results demonstrate 
that the RBF-NN outperformed a standard feed-forward neural network. 
\citeboth{ji2014} also found that the RBF-NN, on a 
small training dataset, was able to predict the value 
of the Shanghai Composite Index on a 45 day time-horizon, with a mean prediction error of 
1.4387$\%$, suggesting that the RBF-NN is competent in time-series 
forecasting using small training datasets. Recently, \citeboth{abotaleb2024}
used quarterly data from 2001 to 2023 to model the influence of economic 
policies on the quarterly prices of various stock markets including the 
FTSE100 using an RBF-NN, finding that the model 
achieved relative forecast error rates as low as 6.8$\%$ and an error of 
27.2$\%$ in the case of the FTSE100. Thus, it is concluded that the RBF-NN is a 
competent tool in the sphere of financial time-series forecasting, and 
is able to operate effectively within the constraints placed by our 
problem domain and data.

@@

Based on the empirical research and characteristics of the UK economy,
the stock market function is defined as:

\begin{align}
    FTSE_{250} = f\biggl(CPI, EXCHG, INT, M3\biggr) \label{eq:implicit}
\end{align}

@@

Before constructing the ARDL model a natural logarithmic transform is applied
on the variables, this is done because in a log-transformed model,
the coefficients can be interpreted as elasticities. Put simply, 
a coefficient in a model, where both the dependent and independent variables
are in logarithmic form, represents the percentage change in the 
dependent variable for a one percent change in an independent variable, 
which furnishes us with easily interpretable results. Having done this,
the level at which the variables are stationary\footnote{Stationarity refers to the statistical property of a time-series where 
its mean, variance, and autocorrelation structure do not change over time. Either a variable is stationary at level ($I(0)$), requiring no alteration,
or it is stationary after taking the first difference ($I(1)$). 
Non-stationarity leads to spurious regression where relationships that 
are not actually present between two time-series are produced.} is checked.
Various techniques exist for stationarity testing, however in this study the 
Augmented Dickey-Fuller (ADF) unit root test, was used.

As the ARDL model involves lagged values of the dependent and independent 
variables, it is necessary to determine the optimum lag for the variables 
that minimises error; this was done using the Schwarz
Information Criterion.

The implicit form \eqref{eq:implicit} is represented in the ARDL model as:

@@

for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic is computed for the joint
significance of the lagged variables and compared against the critical values
given by \citeboth{pesaran2001} to determine whether cointegration exists, 
and at what significance level. If cointegration exists, then 
the long-run coefficients $\kappa_i$ are estimating using the cointegration equation:

\begin{align}
    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \label{eq:two}
\end{align}

@@

where the error correction term ECT is found by calculating the residual 
produced by \eqref{eq:two}:

@@

The parameter $\gamma$ describes the speed at which short-run deviations from the long-run equilibrium
are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that converges the deviation 
to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
well as the short-run correction speed $\gamma$ OLS regression is used.

@@

First, a natural logarithmic transform was applied on the variables, and 
then conducting the ADF unit root test on the data. The null hypothesis of the 
ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
is that the variables are stationary.

@@

Subsequently, the optimum lag of the variables was calculated using the 
Schwarz Information Criterion, to obtain the optimum model 
of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.0798$ was compared against the critical values
given by \citeboth{pesaran2001}, providing evidence of cointegration 
at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
level.

@@

\begin{table}[h!]
    \centering
    \caption{Coefficients Estimates}
    \begin{tabular}{llll}
        \toprule
        \textbf{Variable} & \textbf{Long-Run Coefficient} & \textbf{Variable} & \textbf{Short-Run Coefficient} \\
        \midrule
        FTSE$_{250}$ & N/A & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
        CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
        INT & $-0.1341097^{***}$ & INT $(-7)$ & $-0.034572^{****}$\\
        EXCHG &  $-0.890146^{****}$ & EXCHG $(-7)$ & $0.122318$ \\
        M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.1001^{**}$ \\
        \bottomrule
    \end{tabular}
\end{table}

@@

The long term coefficients show that the interest rate and exchange rate
interest rate have significant long term effects on the stock market price
while money supply and inflation are statistically insignificant. \citeboth{neifar2023}
proposes that this is because deflation, that is the negative movements in the 
CPI, is a statistically significant factor in the movement of the UK stock 
market, but positive movements (inflation) are not. Nevertheless, the positive
coefficient of 3.16, and the negative short-term adjustment, is in line with the expectation from the literature that 
inflation exerts a positive effect on the stock market in the long-term, but a negative 
one in the short-run.

The insignificant long-term effect of the money supply 
was noted by \citeboth{pícha2017} and appears to be confirmed here. There are a few explanations
for why this may be.

Unsurprisingly, the USD/GBP exchange rate is significantly related to the FTSE 250's movements. It 
should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
at times $50\%$ of the FTSE 250's earnings have come from abroad, indicating that 
the value of the Pound has an important impact on this stock market. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it by $1\%$ has an impact of $-0.89\%$ on the FTSE 250. The findings 
are in line with \citeboth{neifar2023} and \citeboth{khan2018}.

@@

Turning to the short-run dynamics, the error correction term 
is negative and statistically significant ensuring that long-run equilibrium 
can be achieved. The adjustment speed of $0.1001$ indicates that 
roughly $10\%$ of the fluctuation in the previous month's price caused by the 
macroeconomic variable movements, is adjusted back to the long-run equilibrium in the current month.
However, only increases in the interest rate and money supply are able to 
significantly contribute towards this correction, with $1\%$ increases in the former 
leading to a $-0.03\%$ change in the FTSE250 price 7-months later, and in the case of the latter
a $-0.53\%$ change after a 4-month lag.

@@

By introducing several modifications to the exact interpolation 
procedure, the RBF-NN
model is obtained, as outlined originally in \citeboth{broomhead1988}. 
The RBF-NN mapping takes the following form:

@@

The RBF-NN was utilised for both short-term and long-term 
predictions. For short-term forecasting, 
the model was tested over a 5-month horizon, based on the optimum lag calculated earlier, while the long-term predictions spanned 57 months. 
The data was differenced based on the recommendations of the 
unit root tests conducted earlier to achieve stationarity.

First, the data was normalised using a z-score normalisation. This is because
neural networks work best with data
that is between a specified range; it ensures that the data is roughly uniformly
distributed between the network inputs and outputs 
(Mendelsohn, 1993, Preprocessing Data for Neural Networks). Following this 
it was necessary to program the RBF-NN into Python. As no specific RBF package exists
in Python, the kernel regression approach outlined by 
\citeboth{bishop1995}, using the GaussianProcessRegressor (GPR) package in Python.

@@

As this is a Bayesian model a \textit{prior distribution} over the parameters must be specified, 
expressing our beliefs about the prior
parameters before the observations are used. A covariance function or kernel
is specified according to the characteristics of the data; in our model this 
is the generalised inverse multiquadric

@@

\begin{align}
    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
\end{align}
where the covariance matrix $A^{-1} = K^{-1} + \frac{XX^T}{\sigma_{n}^2}$.
To make predictions on a test case all possible $w$
values, weighted by their posterior probability, are averaged to obtain the
predictive distribution, upon which predicted outputs are computed. \citep{rasmussen2006}

@@

A $4$-dimensional input space $X$
passes inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \), through the first
layer, in which the covariance function $(x,x')$
is applied, the goal being to determine the distribution of the weights. Prior to this,
the optimal hyperparameters in the covariance function, which plays an analogous
role to the RBF basis function, are calculated using GridSearch Cross Validation; these hyperparameters are:
\begin{enumerate}
    \item $\alpha$: a regularisation parameter which represents a small quantity
    that is added to the diagonals of the covariance matrix to ensure it is 
    invertible. 
    \item $\beta$: the power in the covariance function, which controls the shape of the function.
    \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for; naturally, 
    it moderates the risk of overfitting and underfitting.
    \item Length scale: a parameter that controls the influence - analogous to 
    the width each RBF neuron possesses - that each kernel (analogous to node) is to have.
\end{enumerate}
following this the model uses the training observation $(y,X)$ to find
the posterior predictive distribution, and with it the predicted output, after 
averaging over all possible $w$, weighted by their likelihood according to \eqref{eq:nine}.

@@

The interpretation of this is that, over the long-run (4-5 years), the variations in the 
Consumer Price Index, interest rate, USD/GBP exchange rate, and M3 money 
supply explain or account for $9\%$ of the fluctuations of the
FTSE250 value. This relates closely to the findings of \citeboth{conover1999}, 
who used an econometric model to establish that $4\%$ of the variation in monthly UK stock returns can be 
explained by variations in the money supply.

@@

In the context of a phenomenon as complex as the stock market,
which is influenced by an inumerable quantity of factors such as investor sentiment
- inherently irrational, 
political instability, geopolitical shocks, corporate governance, and financial regulation, this result 
unsurprising. Furthermore, the selection of economic variables in this study
is by no means exhaustive; commodity prices, international trade dynamics,
economic growth, and many other variables also exert a non-insignificant 
effect on the stock market.

@@

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$, for a 5-month time horizon.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possesses
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

@@

In the short-term, which defined as 5 months, 
as per the optimum lag for the FTSE250 in the ARDL model, the market movements
are more closely tied to recent deviations in the economy, because theoretically
the short-run is not exposed to the various unforeseen external shocks and structural changes
that dilute the relationship between these variables and the stock market in the 
long-run. Essentially, as the prediction window expands, 
more uncertainty and unknowns are introduced, which translates into 
lower predictive quality. Furthermore, as the time horizon extends into the long-run, fundamental factors 
begin to matter a lot more. The compound effect of earnings, cash flows and
managerial efficiencies begins to outweigh the 
influence of fluctuating macroeconomic sentiment. Thus, in the short-run 
a different phenomenon is observed.

@@

The the study also investigated the efficacy of machine learning methods in 
the field of stock-econometric forecasting. The RBF-NN performed well in
short-term forecasting of the FTSE250, and offers investors with an accurate forecasting tool, 
to leverage in their trading strategies. The long-term forecast was able to improve on the ARDL-ECM model, and provides 
economists with a tool to compute non-econometric, quantitative measures of the 
long-term influence of macroeconomic variables on the stock market.

Further research should endeavour to include other macroeconomic variables such as 
foreign direct and portfolio investment, commodity prices and economic growth.
Falling outside the immediate scope of the current study, 
an autoregresive model could also be tried, including the past values of the 
FTSE250 in the input space to provide a more accurate model. In addition, 
a more granular data set could be utilised by applying a 
Brownian bridge algorithm to generate daily macroeconomic data and 
subsequently testing the RBF-NN against other neural networks 
such as auto-regressive LSTMs, as well as other machine learning methods suited for 
forecasting such as Support Vector Machines (SVM).
